---
title: "Wikistats exploratory analysis"
author: "Fernando Reis"
date: "20/08/2015"
output: html_document
---

This report presents an exploratory data analysis of the number of page views of articles in the english wikipedia related to World Heritage Sites inscribed in UNESCO official list.

The R markdown source code of this report is available from the GitHub repository:
https://github.com/avirgillito/unece-sandbox2015-wikistat

# Preparation for the data analysis

## Loading libraries and auxiliar code

Required libraries were loaded and auxiliar functions, declared in 'whs_aux.R', were sourced.
```{r message=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(maps)
source("../scripts/r/whs_aux.R")
```

## Loading list of WHS

UNESCO official list was loaded and the list of sites was restricted to those inscribed before 2015.
```{r}
whs <- read.csv("../data/whs.csv", fileEncoding="UTF-8") %>%
	mutate(whs_id = as.numeric(as.character(id_number))) %>%
	filter(date_inscribed < 2015)
```

A summary of the main characteristics of the WHS.
```{r}
summary(whs[, -c(6:15,17:20,23)])
```

And the evolution of nomination of sites over time.
```{r fig.width=10}
plot(table(whs$date_inscribed), type="l")
```

A map of the WHS.
```{r fig.width=10, fig.height=5}
world <- map_data("world")
ggplot(world, aes(long, lat)) + 
	geom_polygon(aes(group = group), fill = "white", color = "gray40", size = .2) + 
	geom_point(data = whs, aes(longitude, latitude, color = category), alpha=0.8, size = 3)
```


## Loading WHS articles

Get list of WHS articles in english wikipedia, deleting commas from the corresponding titles.
```{r}
fileName <- paste0(".", DATA_FOLDER, "/whsArticles.csv")
whsArticles <- read.csv(fileName, fileEncoding="UTF-8") %>%
	mutate(article = gsub(",", "", article))
```

Check number of articles by WHS.
```{r}
t <- as.data.frame(table(whsArticles$whs_id))
table(t$Freq)
```

The highest number of articles is `r max(t$Freq)`, for WHS with id number `r ` '971' which corresponds to 'Churches of Chiloé'.
The 17 articles correspond to each one of the churches.
```{r}
maxFreq <- max(t$Freq)
filter(t, Freq==maxFreq)
filter(whs, id_number==971)$site
filter(whsArticles, whs_id==971)
```

Get articles with WHS ids which are not in the list. These artcles should refer only to sites inscribed after 2014.

*__Note:__ 'id_number' needs to be first converted to character because it is a factor variable.*
```{r}
whsArticles %>%
	left_join(whs[, c("whs_id", "site")]) %>%
	filter(is.na(site)) %>%
	select(article)
```

Get WHS in UNESCO list which do not have articles in the English wikipedia.
```{r}
whs %>%	
	left_join(whsArticles) %>%
	filter(is.na(X)) %>%
	select(site)
```

## Loading WHS articles page views data

Load page views time-series for articles.
```{r}
fileName <- paste0(".", DATA_FOLDER, "/wikistats_en.txt")
whsArtViews <- read.table(fileName, header = T)
whsArtViews$article <- sapply(as.character(whsArtViews$article), FUN=URLdecode)
```

Melt page views data frame.
```{r}
whsArtViews <- melt(whsArtViews, id.vars="article", variable.name="month")
```

Compute total pageviews by article.
```{r}
whsArtViews <- group_by(whsArtViews, article)
aggrArticles <- summarise(whsArtViews, tot_pageviews=sum(value))
aggrArticles <- arrange(aggrArticles, desc(tot_pageviews))
```

Get top 20 articles with highest number of page views.
```{r fig.width=10}
top20 <- aggrArticles[1:20,]
top20$article <- as.character(top20$article)
top20 <- transform(top20, article = reorder(article, tot_pageviews))
ggplot(data=top20, aes(x=article, y=tot_pageviews)) +
        geom_bar(stat="identity") + coord_flip() + scale_y_continuous('') + scale_x_discrete('')
```

Get WHS articles without page views time-series. These normally refer to articles which did not exist during the time reference of the page views time-series used.
```{r}
whsArticles %>%
	left_join(aggrArticles) %>%
	filter(is.na(tot_pageviews)) %>%
	select(article)
```

Get WHS pageviews time-series for articles which are not listed. Normally this should not return any article.
```{r}
aggrArticles %>%
	left_join(whsArticles) %>%
	filter(is.na(X)) %>%
	select(article)
```

# Exploratory data analysis

Page views were aggregated by WHS.
```{r}
whsSiteViews <- whsArtViews %>%
	left_join(whsArticles) %>%
	group_by(whs_id, month) %>%
	summarise(tot_pageviews=sum(value))
```

## Popularity of WHS in Wikipedia

Most popular WHS.
```{r fig.width=10}
popWhs <- whsSiteViews %>%
	group_by(whs_id) %>%
	summarise(tot_pageviews = sum(tot_pageviews)) %>%
	arrange(desc(tot_pageviews)) %>%
	left_join(whs[, c("whs_id", "site")], by="whs_id")

top20 <- popWhs %>%
	slice(1:20) %>%
	mutate(site = strtrim(as.character(site), 80)) %>%
	mutate(site = reorder(site, tot_pageviews))

ggplot(data=top20, aes(x=site, y=tot_pageviews)) +
        geom_bar(stat="identity") + coord_flip() + scale_y_continuous('') + scale_x_discrete('')
```

Number of WHS per country.
```{r fig.width=10}
cntrSites <- whs %>%
	select(whs_id, transboundary, iso_code) %>%
	filter(transboundary == 0) %>%
	group_by(iso_code) %>%
	tally() %>%
	arrange(desc(n)) %>%
	slice(1:20) %>%
	mutate(iso_code = reorder(iso_code, desc(n)))

ggplot(data = cntrSites, aes(x = iso_code, y = n)) + 
	geom_bar(stat = "identity")
```

Popularity of WHS per country.
```{r fig.width=10}
cntrViews <- whs %>%
	select(whs_id, transboundary, iso_code) %>%
	filter(transboundary == 0) %>%
	left_join(whsSiteViews) %>%
	select(iso_code, tot_pageviews) %>%
	replace(is.na(.), 0) %>%
	group_by(iso_code) %>%
	tally(wt = tot_pageviews) %>%
	arrange(desc(n)) %>%
	slice(1:20) %>%
	mutate(iso_code = reorder(iso_code, desc(n)))

ggplot(data = cntrViews, aes(x = iso_code, y = n)) + 
	geom_bar(stat="identity")
```

A map of the WHS popularity.
```{r fig.width=10, fig.height=5}
mapDat <- popWhs %>%
	left_join(whs[, c("whs_id", "category", "latitude", "longitude")])
	
world  = map_data("world")
ggplot(world, aes(long, lat)) + 
	geom_polygon(aes(group = group), fill = "white", color = "gray40", size = .2) + 
	geom_point(data = mapDat, aes(longitude, latitude, color = category, alpha=log(tot_pageviews)), size = 3)
```

Popularity of WHS per category. On average mixed sites are more popular.
```{r}
catViews <- whs %>%
	select(whs_id, category) %>%
	left_join(whsSiteViews) %>%
	select(category, tot_pageviews) %>%
	replace(is.na(.), 0) %>%
	group_by(category) %>%
	summarise(avg_pageviews = mean(tot_pageviews)) %>%
	arrange(desc(avg_pageviews)) %>%
	mutate(category = reorder(category, desc(avg_pageviews)))

ggplot(data = catViews, aes(x = category, y = avg_pageviews)) + 
	geom_bar(stat="identity")

whs %>%
	select(whs_id, category) %>%
	group_by(category) %>%
	tally()
```

The most popular Mixed sites.
```{r}
whsMixViews <- whs %>% 
	left_join(whsSiteViews) %>% 
	filter(category=="Mixed") %>% 
	transform(site=paste0(site, " (", iso_code, ")")) %>%
	select(site, tot_pageviews) %>%
	replace(is.na(.), 0) %>%
	group_by(site) %>% 
	summarise(tot_pageviews=sum(tot_pageviews))

top20Mix <- whsMixViews %>%
	arrange(desc(tot_pageviews)) %>%
	transform(site=reorder(site, tot_pageviews)) %>%
	slice(1:20)

ggplot(data=top20Mix, aes(x=site, y=tot_pageviews)) +
        geom_bar(stat="identity") + 
	coord_flip() + 
	scale_y_continuous('') + 
	scale_x_discrete('') +
	ggtitle("Total number of page views during 2012-2013 for the 20 most popular mixed sites")
```


```{r}
whsViews <- whs %>% 
	left_join(whsSiteViews) %>% 
	select(site, category, tot_pageviews) %>%
	replace(is.na(.), 0) %>%
	group_by(site, category) %>% 
	summarise(tot_pageviews=sum(tot_pageviews))

ggplot(whsViews, aes(x=category, y=log(tot_pageviews))) + 
          geom_point(colour="blue", alpha=0.5, position="jitter") + 
          geom_boxplot(outlier.size=3, alpha=0.2)

ggplot(whsViews, aes(x=log(tot_pageviews), group=category, fill=category)) + 
	geom_density(alpha=0.5)

```

Popularity of WHS by year inscribed.
```{r}
inscViews <- whs %>%
	select(whs_id, date_inscribed) %>%
	left_join(whsSiteViews) %>%
	select(date_inscribed, tot_pageviews) %>%
	replace(is.na(.), 0) %>%
	group_by(date_inscribed) %>%
	summarise(avg_pageviews = mean(tot_pageviews))

ggplot(data = inscViews, aes(x = date_inscribed, y = avg_pageviews)) + 
	geom_line(stat="identity") +
	geom_point(stat="identity") +
	ggtitle("Average number of page views during 2012-2013 according to the date of inscription")

inscViews <- whs %>%
	select(whs_id, date_inscribed) %>%
	left_join(whsSiteViews) %>%
	select(date_inscribed, tot_pageviews) %>%
	replace(is.na(.), 0) %>%
	group_by(date_inscribed) %>%
	summarise(tot_pageviews = sum(tot_pageviews))

ggplot(data = inscViews, aes(x = date_inscribed, y = tot_pageviews)) + 
	geom_line(stat="identity") +
	geom_point(stat="identity") +
	ggtitle("Total number of page views during 2012-2013 according to the date of inscription")
```


## Popularity of WHS over time

```{r}
timeWhs <- whsSiteViews %>%
	group_by(month) %>%
	summarise(tot_pageviews=sum(tot_pageviews))

tmPlot <- ggplot(data=timeWhs, aes(x=month, y=tot_pageviews, group=1)) + 
	expand_limits(y=0) +
	geom_line(size=1.6) + 
	geom_point(colour="red", size=4, shape=21, fill="white") + 
	xlab("Month") + ylab("Total number of page views") +
	ggtitle("Page views of English Wikipedia articles related to World Heritage Sites")
tmPlot
```

What about the most popular site?
```{r}
top20Time <- whsSiteViews %>%
	inner_join(top20, by="whs_id") %>%
	mutate(site = substr(site, 1, 30)) %>%
	transform(site=reorder(site, tot_pageviews.y))
	

tmPlot + geom_area(data=top20Time, 
		   aes(x=month, y=tot_pageviews.x, group=site, fill=site, order = desc(site)), 
		   colour = 1)
```

WHS which grew the most from 2012 and 2013.
```{r}
gro <- whsSiteViews %>%
	mutate(year=paste0("y", substr(month, 2, 5))) %>%
	group_by(whs_id, year) %>%
	summarise(tot_pageviews=sum(tot_pageviews)) %>%
	dcast(whs_id ~ year) %>%
	mutate(growth = y2013 / y2012 - 1)

groTop20 <- gro %>%
	left_join(whs, by="whs_id") %>%
	select(whs_id, site, iso_code, y2013, y2012, growth) %>%
	filter(y2012 != 0, y2013 != 0) %>%
	arrange(desc(growth)) %>%
	transform(site=reorder(site, growth)) %>%
	slice(1:20) %>%
	mutate(site=paste0(substr(site, 1, 40), " (", iso_code, ")"))

groTop20[, c("site", "growth")]
```

WHS which decreased the most from 2012 and 2013.
```{r}
groBottom20 <- gro %>%
	left_join(whs, by="whs_id") %>%
	select(whs_id, site, iso_code, y2013, y2012, growth) %>%
	filter(y2012 != 0, y2013 != 0) %>%
	arrange(growth) %>%
	transform(site=reorder(site, growth)) %>%
	slice(1:20) %>%
	mutate(site=paste0(substr(site, 1, 40), " (", iso_code, ")"))

groBottom20[, c("site", "growth")]
```

